---
title: "Class 5: Data Viz with ggplot"
author: "Steven (PID:A14837063)"
format: pdf
---

## Running Code
A very popular package in this area is called **ggplot2**

Before I can use any add-on package. I must install it with the `install.packages("ggplot2')` command/function

Then to use the package I need to load it with a `library(ggplot2)` call.

```{r}
library(ggplot2)
ggplot(cars) +
  aes(x=speed, y=dist)+
  geom_point()
```

For "simple" plots like this one base R code will be much shorter than ggplot code. 
```{r}
ggplot(cars) +
  aes(x=speed, y=dist)+
  geom_point() +
  geom_smooth() 
```

Every ggplot has at least 3 layers

-**data** (data.frame with the numbers and stuff you want to plot)
-**aes**thetics (mapping of your data columns to your plot)
-**geom**s (there are tones of these, basics are `geom_point()`, `geom_line()`, `geom_cols()`)

```{r}
mtcars
ggplot(mtcars) +
  aes(x=mpg, y=disp)+
  geom_point() 

```
```{r}
mtcars
ggplot(mtcars) +
  aes(x=mpg, y=disp, size= hp)+
  geom_point(color='blue') 
```
```{r}
library(ggrepel)
ggplot(mtcars) +
  aes(x=mpg, y=disp, col = am, label=rownames(mtcars))+
  geom_point() +
  facet_wrap(~am)
  geom_text_repel

```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
p <- ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point()
p

p + scale_colour_manual(values=c("blue","gray","red")) +
    labs(title="Gene Expresion Changes Upon Drug Treatment",
         x="Control (no drug) ",
         y="Drug Treatment")
```
There are `nrow(genes)` in this dataset
```{r}
library(gapminder)
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.5)
```
```{r}
library(gapminder)
gapminder_1957 <- gapminder%>% filter(year==1957)
ggplot(gapminder_1957) + 
  aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 10) 
```
```{r}
library(gapminder)
gapminder_2007 <- gapminder%>% filter(year==2007)
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 10) 
```
The `table` is a super useful utility to tell me how many entries of each type there are
Key points: 
 saving plots with ggsave
 different plots types with different geoms()
 Fcacetiwng with 
 Multi-plot layout with the **patchwork**
 
 p1 |p2|p3 /p4
```{r}
ggplot(mtcars) +
  aes(mpg,disp)+
  geom_point()
```
the functions `nrow()`, `ncol()`, and `table()` are ones I want you to know. 